---
title: "Fitting Pythagorean Model in NBA"
author: "Chloe Isabella Tsang"
date: "19/08/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(stringr)
library(rvest)
library(xml2)
```

## Reparameterization 

We are fitting a pythagorean model for NBA games in January, 2019. 

Given Pythagorean model: 
Win%=(PW^a)/((PW^a)+(PL^a))

Need to reparemeterize the equation to get it into linear form. We do it as so:  

Win%/(1-Win%)=(PW/PL)^a

Take log of both sides to bring the a exponent down as a co-efficient.

log(Win%/(1-Win%))=alog(PW/PL)  

In this form of y=mx:
y=log(Win%/(1-Win%)), slope(m)=a, x=log(PW/PL)

Next, use the regression function glm() with y and x above.  

## Using the glm() function

Previously, nba_scores function has been used to gather data from specified url. 

```{r echo=TRUE}
nba_scores<-function(url){
data<-read_html(url) %>%
html_nodes("table") %>%
html_table()
data<-data[[1]]
data<-data[,c(1,3:6)]
names(data)<-c("Date","Visitor","Visitor.pts","Home","Home.pts")
rbind(data %>% select(Date,Team=Visitor, Opponent=Home, Pts=Visitor.pts,Opponentspts=Home.pts),data %>% select(Date, Team=Home, Opponent=Visitor, Pts=Home.pts, Opponentspts=Visitor.pts))
}
```
Using a URL, and inputting it into the function:
```{r echo=TRUE}
data<-nba_scores("https://www.basketball-reference.com/leagues/NBA_2019_games-january.html") 
```
Replace this 'data' object with another object (eg. plotdata). This new object will summarise data as so: 
```{r echo=TRUE}
plotdata<-data %>%
  group_by(Team) %>%
  dplyr::summarise(
    PtsWon=sum(Pts),
    PtsLost=sum(Opponentspts),
    Win=mean(Pts>Opponentspts)
  ) 
```

The glm function has the following syntax: glm(formula,data,family) - where formula is inputted though y~x.

Earlier in the reperameterization, we found y and x through the linear form of the logistic model. Here, use the earlier x and y, enter:
```{glm input, echo=TRUE}
glm(log((Win)/(1-Win))~log(PtsWon/PtsLost),data=plotdata,family=gaussian)
```
This should give us details regarding the model, including coefficients('intercept' and 'log(PtsWon/PtsLost)'). We can add -1 to the x component of 'formula' as to get rid of intercept. 
```{glm w/o intercept, echo=TRUE}
glm(log((Win)/(1-Win))~log(PtsWon/PtsLost)-1,data=plotdata,family=gaussian)
```
The value you get for coefficient 'log(PtsWon/PtsLost)' should be the slope of the model, as well as your value for alpha (a).

Use plotdata and the pipe function to get the plot:
```{r echo=TRUE}
plotdata%>%
ggplot(aes(y=log((Win)/(1-Win)),x=log(PtsWon/PtsLost)))+geom_point(size=2,col="#1792d0") +
geom_smooth(method = "lm")
```

using data from "https://www.basketball-reference.com/leagues/NBA_2019_games-january.html" 

## My interpretation of the corresponding changes made to alpha relating to model
The reparameterization was done to fit a linear model, because the original form of the pythagorean model won't work. After doing this, you manage to get the explanatory variables condensed in a way that you can replace them with another substitute variable that still fits 2 variables for model axes - one as x, the other as y. 

The PW and PL variables can be substituted by X1, for example.

log(Win%/1-Win%)=a(logX1), with X1=PW/PL

So, x=log(Win%/1-Win%), y=log(X1), leaving alpha (a) as the slope*.

*Slope of a regression describes the predicted values of y given x, the rate of the change in y as x changes, since y is dependent on x.

## Looking at different alpha values for different situations
Over-look of alpha value:
```{r echo=TRUE}
data<-nba_scores("https://www.basketball-reference.com/leagues/NBA_2019_games-january.html")
plotdata<-data %>%
  group_by(Team) %>%
  dplyr::summarise(
    PtsWon=sum(Pts),
    PtsLost=sum(Opponentspts),
    Win=mean(Pts>Opponentspts)
  )
glm(log((Win)/(1-Win))~log(PtsWon/PtsLost)-1,data=plotdata,family=gaussian)
```

The 15.56 value here is the alpha found from this version of the fitted pythagorean model for our used data. 








