---
title: "Shrinkage (part 4)"
author: "Chloe Isabella Tsang"
date: "8/28/2019"
output: html_document
---

```{r setup, include=FALSE}
library(deuce)
library(dplyr)
library(tidyr)
library(ggthemes)
library(ggplot2)
library(scales)
library(htmlTable)
library(stringr)
library(lubridate)
library(lme4)
```

## Learning Hierarchies
Hierarchies matter because 'grouped' observations tend to correlate, but standard models treat observations as independent - need special models to accound for dependence in hierarchial data. 

## Looking at explample w Service Data

Good service model: can help set outcome expectation of a player's serve, determine if player palyed or worse than expectation, identitfy factors that influence serve performance.

Loading stats organized by match winner and loser:
```{r, echo=TRUE, eval=TRUE}
data(atp_matches) 
tail(atp_matches %>%
filter(tourney_level == "Grand Slams") %>%
select(tourney_name, surface, winner_name, loser_name, w_svpt, w_1stWon, w_2ndWon))
```

Prepare men's service outcomes for grand slams in 2017:
```{r prep (1)}
slams_2017 <- atp_matches %>%
    filter(year == 2017, tourney_level == "Grand Slams") %>%
    select(tourney_name, surface, 
           winner_name, w_svpt, w_1stWon, w_2ndWon,
           loser_name, l_svpt, l_1stWon, l_2ndWon) %>%
    dplyr::mutate(
      w_servewon = w_1stWon + w_2ndWon,
      l_servewon = l_1stWon + l_2ndWon
    )

winner <- loser <- slams_2017
names(winner) <- sub("winner|w_", "player", names(winner))
names(loser) <- sub("loser|l_", "player", names(loser))
names(loser) <- sub("winner|w_", "opponent", names(loser))
names(winner) <- sub("loser|l_", "opponent", names(winner))
slams_2017 <- rbind(winner, loser)


slams_2017 %>%
  group_by(player_name) %>%
  dplyr::summarise(
    matches = n()
  ) %>%
  ggplot(aes(x = player_name, y = matches)) + 
  geom_bar(stat = "identity", fill = "#e5952c") + 
  coord_flip() + 
  theme_hc() + 
  scale_y_continuous("Matches") + 
  scale_x_discrete("")
```

*Given data above, need to consider player average serve ability, opponent's average return ability, and surface.  

*Look if there are any hierarchies? 
Yes, there are because there are different matches and service points across the players.  

Fitting hierarchial model has 2 approaches - Likelihood-based, and Bayesian. 

Use the lme4 packages for fitting hierarchial models for exponenetial family distributions. Here, is a Likelihood based binomial logistic model for serve win proportion:
```{r binomial logistic model, echo=TRUE}
fit <- glmer(
  cbind(playerservewon, playersvpt - playerservewon) ~ 
    I(surface == "Grass") +
  I(surface == "Clay") +
    (1 | player_name) +
    (1 | opponent_name),
  data=slams_2017, family="binomial"
)
fit
```
Notes:
- the (1 | group) is normal random effects for a 'group' variable. 
- use binomial as default for logit link - (family="binomial" is different from family="gaussian")

We can get a summary of the model with the summary function:
```{r summary, eval=TRUE, echo=TRUE}
summary(fit)
```

